---
title: "A simple workflow for PoweREST"
author: "Lan Shui"
date: "2024-06-27" # "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PoweREST}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette highlights a simple example workflow for performing power analysis for ST data using the `PoweREST` R package. A detailed version can be found on PoweREST GitHub page.

## Load Packages
Once installed, **`PoweREST`** can be simply loaded (along with the required packages) as follows:
```{r setup}
#install.packages("devtools")
#devtools::install_github("lanshui1998/PoweREST")
#----or
#install.packages("PoweREST")
library(PoweREST)
```

## Prepare data
```{r,eval=FALSE}
#load ST data in R by Seurat:
#here we load the pancreatic cancer data which is available on GitHub page
three_areas <- readRDS("your path to/GSE233293_scMC.all.3areas.final")
Idents(three_areas)
#Levels: Peri Juxta Epi
SeuratObject_splitlist<-Seurat::SplitObject(three_areas, split.by = "ident")

#split the ST data into three areas
for (i in 1:length(SeuratObject_splitlist)) {
  SeuratObject_splitlist[[i]][['Condition']]<-ifelse(SeuratObject_splitlist[[i]][['Type']]=='LG','LG','HR')
}

for (i in 1:length(SeuratObject_splitlist)) {
  Seurat::Idents(SeuratObject_splitlist[[i]])<-"Condition"
}

# Take Peri area for example for downstream analysis
Peri<-SeuratObject_splitlist$Peri
```

## Compute power values through Bootstrapping
```{r,eval=FALSE}
result<-PoweREST(Peri,cond='Condition',replicates=5,spots_num=80,iteration=100)

#---For test, try this first
#PoweREST(Peri,cond='Condition',replicates=5,spots_num=80,iteration=2)
#---To get faster, try this
#devtools::install_github('immunogenomics/presto')
```

### Instead of using default Wilcoxon test
```{r,eval=FALSE}
# For example, use the Student's t-test
result2<-PoweREST(Peri,cond='Condition',replicates=5,spots_num=80,iteration=100,test.use="t")
```

Users can also use PoweREST_gene and PoweREST_subset to perform the power estimation upon one gene or a subset of genes.
### PoweREST_gene
```{r,eval=FALSE}
PoweREST_gene(Peri,cond='Condition',replicates=5,spots_num=80,gene_name='MUC1',pvalue=0.00001)
```

### PoweREST_subset
```{r,eval=FALSE}
PoweREST_subset(Peri,cond='Condition',replicates=5,spots_num=80,pvalue=0.05,logfc.threshold = 0.1,min.pct = 0.01)
```

## Fit the power surface using smoothing splines
```{r,eval=FALSE}
#Fit the power surface for sample size=5 in each arm
b<-fit_powerest(result$power,result$avg_logFC,result$avg_PCT)
```

### Visualize the power surface
```{r,eval=FALSE}
pred <- pred_powerest(b,xlim= c(0,6),ylim=c(0,1))
vis_powerest(pred,theta=-30,phi=30,color='heat',ticktype = "detailed",xlim=c(0,6),nticks=5)
```

### Create interactive visualization result
```{r,eval=FALSE}
plotly_powerest(pred,fig_title='Power estimation result')
```

## Fit local power surface with XGBoost
```{r,eval=FALSE}
# Fit the local power surface of avg_log2FC_abs between 1 and 2
avg_log2FC_abs_1_2<-dplyr::filter(power,avg_log2FC_abs>1 & avg_log2FC_abs<2)
# Fit the model
bst<-fit_XGBoost(power$power,avg_log2FC=power$avg_log2FC_abs,avg_PCT=power$mean_pct,replicates=power$sample_size)
# Make predictions
pred<-pred_XGBoost(bst,n.grid=30,xlim=c(0,1.5),ylim=c(0,0.1),replicates=3)
```

### Visualize the local power surface
```{r,eval=FALSE}
#2D version
vis_XGBoost(pred,view='2D',legend_name='Power',xlab='avg_log2FC_abs',ylab='mean_pct')
#3D version
vis_XGBoost(pred,view='3D',legend_name='Power',xlab='avg_log2FC_abs',ylab='mean_pct')
```
